Discussion Draft

### 2020-04-15 doc/Discussion.rmd
cleaning up the writing

other COM, / nucleus membrane citations (Harper, Golubovskaya, & Cande, 2004; Naranjo & Corredor, 2008; Scherthan et al., 1996; Zickler & Kleckner, 2016).

telomere guided movements  citations: (Bass et al.,
2000; Chacon, Delivani, & Tolic, 2016; Curtis, Lukaszewski, & Chrzastek, 1991; Ding, Yamamoto, Haraguchi, & Hiraoka, 2004; Gerton &
Hawley, 2005; Lee, Conrad, & Dresser, 2012; Lefrancois, Rockmill,
Xie, Roeder, & Snyder, 2016; Page & Hawley, 2003)

Repair of DSB - CO
(Anderson & Stack, 2005; Bass et al., 2000; Brown et al., 2005; Croft & Jones, 1989; Higgins, Osman, Jones, & Franklin, 2014; Klutstein & Cooper, 2014, Lukaszewski, 1997; Pratto et al., 2014; Viera, Santos, & Rufas, 2009;
Xiang, Miller, Ross, Alvarado, & Hawley, 2014).

@@

Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment. <We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle before starting the irreversible metaphase-to-anaphase transition (cite).  A stable chromosome structure, known as bi-orientation, requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion (cite). Unstable bivalents on the meiosis I spindle, cause the SAC to delay anaphase and allow time for the bivalent to stabilize on the microtubule spindle. If they fail to attach, the SAC triggers apoptosis. Genetic variants which increase the number of unstable bivalents (such as those affecting recombination rate or cohesion), will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle. Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis (LaneKauppi, Subrahman). 
These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volume (cite). This model nominates conserved differences in in the context of meiosis between males and females as the primary source of distinct evolutionary trajectories in genome wide recombination rates.

Regardless of if these hypotheses turn out to be true, our results extend previous reports (dumont) of rapid male-specific evolution in multiple subspecies and strains of house mouse.

whether these instances of evolution are due to seperat independent events or segregating genetic variation within the population requires further study.

These results may hold importance to animal breeding programs -- may suggest that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most efficient way to increase the sex-averaged recombination rates. This could be important application for animal breeding programs (Battagin et al 2016).

@@

Our results of greater between cell variation in females are supported by empiracal results from human and mouse [@lynn2002, @gruhn2013].

>>>>

@@

Under the spindle based model, faster male evolution of the genome wide recombination rate, is primarily driven by a more efficient spindle assembly checkpoint (SAC) acting on chromosome structures at metaphase I alignment. <We refer the reader to [@laneKauppi2019] for more details on the SAC pathway.> The SAC prevents anueploidy by ensuring all bivalents are correctly attached to the microtubulue spindle before starting the irreversible metaphase-to-anaphase transition (cite).  A stable chromosome structure, through bi-orientation, requires, i) kinetochores attached to opposite poles of the cells and ii) at least one crossover to create tension across a proportion of sister cohesion (cite). Unstable bivalents on the meiosis I spindle, cause the SAC to delay anaphase and allow time for the bivalent to stabilize on the microtubule spindle. If they fail to attach, the SAC triggers apoptosis. Genetic variants which increase the number of unstable bivalents (such as those affecting recombination rate or cohesion), will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle. Multiple lines of evidence indicate that the effectiveness of SAC in spermatogenesis is greater compared to oogenesis (LaneKauppi, Subrahman). 
These differences in SAC strength have been connected to conserved features of gametogenesis; centrosome spindle and cell volume (cite). This model nominates conserved differences in in the context of meiosis between males and females as the primary source of distinct evolutionary trajectories in genome wide recombination rates.

Regardless of if these hypotheses turn out to be true, our results extend previous reports [@dumont2011] of rapid male-specific evolution in multiple subspecies and strains of house mouse. Whether these instances of evolution are due to separate independent events or segregating genetic variation within the population requires further study. The faster male evolution in genome-wide recombination rates suggests that even for species with female biased heterochiasmy, selecting for higher male specific genome wide recombination rates may be the most effective way to increase the sex-averaged recombination rate for a genetic background. This insight could be important for animal breeding programs [@battagin2016].

@@

Our results of greater between cell variation in females are supported by empirical results from human and mouse [@lynn2002, @gruhn2013].
